### 2020-05-01 final_report/results.tex
Minor edits
fix typesetting again

We begin by comparing a diverse set of surrogate families that we proposed
to evaluate on previously generated samples of the expensive MC TBR model.
Through the four experimental cases described
in~\cref{sec:experiment-methodology}, we aim to study properties of the
considered models in terms of regression performance, training and prediction
time.

@@

In the third experiment we examine surrogate scaling properties by correlating
metrics of interest with training set size. Firstly, the results shown 
in~\cref{fig:scaling-r2} (and listed in~\cref{tbl:exp3-detailed-results-r2} in
the Appendix) suggest that the most accurate families from the previous experiments
consistently maintain their relative advantage over others, even as we introduce
more training points. While such families achieve nearly comparable regression
performance on the largest dataset, in the opposite case methods based on
decision trees clearly outperform neural networks. This can be observed
particularly on sets of sizes up to~\num{6000}.

@@

Finally, we study scaling with respect to the mean prediction time (shown
in~\cref{fig:scaling-pred} and listed in~\cref{tbl:exp3-detailed-results-t-pred} in
the Appendix). Our initial observation is that all tested
surrogate families with the exception of previously mentioned instance-based
learning methods offer desirable characteristics overall. Analogous to previous
experiments, GBTs, ABTs and ANNs appear to be tied, as they not only exhibit
comparable times but also similar scaling slopes. Following that, we notice a
clear hierarchy of ERTs, SVMs, GPR and RBFs, trailed by IDW and KNNs.

@@

Having selected ANNs, GBTs, ERTs, RBFs and SVMs based on the results of
experiments~2-3, we utilised the best-performing hyperparameter
assignments. In pursuit of goal~(a), the best approximator (no.~1,
ANN) achieved~$R^2=\num{0.998}$ and mean prediction
time~$\overline{t}_{\text{pred.}}=\SI{1.124}{\micro\second}$. These correspond
to a standard error of regression~$S=\num{0.013}$ and a relative speedup~$\omega=\num{6916416} \times$
with respect to the MC TBR evaluation baseline measured during run~1
(see~\cref{tbl:sampling-runs} for details). Satisfying
goal~(b), the fastest model (no.~2, ANN) achieves~$R^2=\num{0.985}$,
$\overline{t}_{\text{pred.}}=\SI{0.898}{\micro\second}$, $S=\num{0.033}$
and~$\omega=\num{8659251} \times$.
While these surrogates
were trained on the entire available set of~\num{500000} datapoints, to satisfy goal~(c) we also trained a more simplified model (no.~4, GBT)
that achieved~$R^2=\num{0.913}$,
$\overline{t}_{\text{pred.}}=\SI{6.125}{\micro\second}$, $S=\num{0.072}$ and $\omega=\num{1269777} \times$
with a set of size only~\num{10000}.

\begin{table}[h]
	\centering
	\sisetup{round-mode=places,round-precision=3,detect-weight=true,detect-family=true}
	\setlength\tabcolsep{2pt}
	{\scriptsize
		\begin{tabular}{lrrrrrrrr}
		\toprule
		{} & {} & \multicolumn{4}{c}{Regression performance} &
		\multicolumn{3}{c}{Complexity}\\
		\cmidrule(lr){3-6}
		\cmidrule(lr){7-9}
		Model & $|\mathcal{T}|$ & MAE [TBR] & $S$ [TBR] & $R^2$ [rel.] & $R^2_{\text{adj.}}$ [rel.]
						& $\overline{t}_{\text{trn.}}$ [\si{\milli\second}] &
		$\overline{t}_{\text{pred.}}$ [\si{\milli\second}] & $\omega$ [rel.]\\
		\midrule
		\input{tbl/table11}
		\bottomrule
		\end{tabular}
	}
	\caption{Results of experiment~4. Here, means and
		standard deviations are reported over 5~cross-validation folds. In addition, we also give
		cross-validation set size $|\mathcal{T}|$ (in thousands of datapoints)
		and relative speedup $\omega$ with respect to the
		MC TBR model mean evaluation time \SI{7.777049573054314}{\second} (per
		sample) measured during run~1 (see~\cref{tbl:sampling-runs} for details).
		The best-performing metrics are highlighted in bold.}
	\label{tbl:exp4-detailed-results}
\end{table}

Overall we found that due to their superior performance, boosted tree-based
approaches seem to be advantageous for fast surrogate modelling on relatively small training
sets (up to the order of~$10^4$). Conversely, while neural networks perform
poorly in such a setting, they become dominant on larger training
sets (at least of the order of~$10^5$) both in terms of regression performance
and mean prediction time.

\subsection{Results of Adaptive Sampling}
\label{sec:adaptiveres}
\begin{wrapfigure}[10]{r}{0.42\textwidth}
	\centering
	\vspace{-13ex}
	\includegraphics[width=0.42\textwidth]{fig5_sintoy}
	\caption{Sinusoidal toy TBR theory over two continuous parameters,
	$n=1$.}
	\label{fig:sintoy}
\end{wrapfigure}

In order to test our QASS prototype, several functional toy theories for TBR were developed as alternatives to the expensive MC model. By far the most robust of these was the following sinusoidal theory with adjustable wavenumber parameter $n$:

@@

plotted in~\cref{fig:sintoy} for $n=1$ and two continuous parameters $C$.

ANNs
trained on this model demonstrated similar performance to those on the expensive
MC model. QASS performance was verified by training a $\text{1h3f}(256)$ ANN on
the sinusoidal theory for varied quantities of initial, incremental, and MCMC
candidate samples. Although the scope of this project did not include thorough
searches of this hyperparameter domain, sufficient runs were made to identify
some likely trends.

An increase in initial samples with increment held constant had a strong impact on final surrogate precision, an early confirmation of basic functionality. An increase in MCMC candidate samples was seen to have a positive but very weak effect on final surrogate precision, suggesting that the runtime of MCMC on each iteration can be limited for increased efficiency. The most complex dynamics arose with the adjustment of sample increment, shown in~\cref{fig:qassincr}. For each tested initial sample quantity N, the optimal number of step samples was seen to be well-approximated by $\sqrt{N}$. The plotted error trends suggest that
incremental samples larger than this optimum give slower model improvement on both the training and evaluation sets, and a larger minimum error on the evaluation set. This performance distinction is predicted to be even more significant when trained on the expensive MC model, where the number of sample evaluations will serve as the primary bottleneck for computation time.
\begin{figure}[h!]
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=1.1\linewidth]{fig6a_qassincrsamp.pdf}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=1.1\linewidth]{fig6b_qassincrtime.pdf}
    \end{subfigure}
    \caption{QASS absolute training error over total sample quantity (left) and number of iterations (right). MAE represents surrogate error on the adaptively-sampled training/test set, and E\_MAE on the independent evaluation sets.}
    \label{fig:qassincr}
\end{figure}

@@

\begin{figure}[h]
  \centering
    \includegraphics[width=0.78\linewidth]{fig7_qasssampling.pdf}
    \caption{Absolute training error for QASS, baseline scheme, and mixed
	scheme.}
  \label{fig:qasssampling}
\end{figure}

\begin{figure}[h]
  \centering
    \includegraphics[width=0.78\linewidth]{fig8_qasssampling100k.pdf}
    \caption{Absolute training error for QASS and baseline scheme, with 100k initial samples.}
  \label{fig:qasssampling100k}
\end{figure}

>>>>

We begin by comparing the performance of a diverse set of surrogate families on previously generated samples of the expensive MC TBR model.
Through the four experimental cases described
in~\cref{sec:experiment-methodology}, we aim to study properties of the
considered models in terms of regression performance, training and prediction
time.

@@

In the third experiment we examine surrogate scaling properties by correlating
metrics of interest with training set size. Firstly, the results shown 
in~\cref{fig:scaling-r2} (and listed in~\cref{tbl:exp3-detailed-results-r2} in
the Appendix) suggest that the most accurate families from the previous experiments
consistently maintain their relative advantage over others, even as we introduce
more training points. While such families achieve nearly comparable
performance on the largest dataset, in the opposite case tree-based approaches
clearly outperform ANNs. This can be observed
particularly on sets of sizes up to~\num{6000}.

@@

Finally, we study scaling with respect to the mean prediction time (shown
in~\cref{fig:scaling-pred} and listed in~\cref{tbl:exp3-detailed-results-t-pred} in
the Appendix). Our initial observation is that all tested
families with the exception of previously mentioned instance-based
models offer desirable characteristics overall. Analogous to previous
experiments, GBTs, ABTs and ANNs appear to be tied, as they not only exhibit
comparable times but also similar scaling slopes. Following that, we notice a
clear hierarchy of ERTs, SVMs, GPR and RBFs, trailed by IDW and KNNs.

@@

Having selected ANNs, GBTs, ERTs, RBFs and SVMs based on the results of
experiments~2-3, we utilised the best-performing hyperparameters.
In pursuit of goal~(a), the best approximator (no.~1,
ANN) achieved~$R^2=\num{0.998}$ and mean prediction
time~$\overline{t}_{\text{pred.}}=\SI{1.124}{\micro\second}$. These correspond
to a standard error~$S=\num{0.013}$ and a relative speedup~$\omega=\num{6916416} \times$
with respect to the MC TBR evaluation baseline measured during run~1
(see~\cref{tbl:sampling-runs} for details). Satisfying
goal~(b), the fastest model (no.~2, ANN) achieves~$R^2=\num{0.985}$,
$\overline{t}_{\text{pred.}}=\SI{0.898}{\micro\second}$, $S=\num{0.033}$
and~$\omega=\num{8659251} \times$.
While these surrogates
were trained on the entire available set of~\num{500000} datapoints, to satisfy goal~(c) we also trained a simplified model (no.~4, GBT)
that achieved~$R^2=\num{0.913}$,
$\overline{t}_{\text{pred.}}=\SI{6.125}{\micro\second}$, $S=\num{0.072}$ and $\omega=\num{1269777} \times$
with a set of size only~\num{10000}.

\begin{table}[h]
	\centering
	\sisetup{round-mode=places,round-precision=3,detect-weight=true,detect-family=true}
	\setlength\tabcolsep{2pt}
	{\scriptsize
		\begin{tabular}{lrrrrrrrr}
		\toprule
		{} & {} & \multicolumn{4}{c}{Regression performance} &
		\multicolumn{3}{c}{Complexity}\\
		\cmidrule(lr){3-6}
		\cmidrule(lr){7-9}
		Model & $|\mathcal{T}|$ & MAE [TBR] & $S$ [TBR] & $R^2$ [rel.] & $R^2_{\text{adj.}}$ [rel.]
						& $\overline{t}_{\text{trn.}}$ [\si{\milli\second}] &
		$\overline{t}_{\text{pred.}}$ [\si{\milli\second}] & $\omega$ [rel.]\\
		\midrule
		\input{tbl/table11}
		\bottomrule
		\end{tabular}
	}
	\caption{Results of experiment~4. Here, figures are reported over 5~cross-validation folds,
		$|\mathcal{T}|$~denotes cross-validation set size ($\times 10^3$)
		and $\omega$ is a relative speedup with respect to
		$\overline{t}_{\text{eval.}}=\SI{7.777049573054314}{\second}$
		measured in the MC TBR model during run~1 (see~\cref{tbl:sampling-runs}).
		The best-performing metrics are highlighted in bold.}
	\label{tbl:exp4-detailed-results}
\end{table}

Overall we found that due to their superior performance, boosted tree-based
approaches seem to be advantageous for fast surrogate modelling on relatively small training
sets (up to the order of~$10^4$). Conversely, while ANNs perform
poorly in such a setting, they dominate on larger training sets (at least of the
order of~$10^5$) both in terms of regression performance and mean prediction time.

\begin{wrapfigure}[10]{r}{0.4\textwidth}
	\centering
	\vspace{-16ex}
	\includegraphics[width=\linewidth]{fig5_sintoy}
	\caption{Sinusoidal toy TBR theory over two continuous parameters,
	$n=1$.}
	\label{fig:sintoy}
\end{wrapfigure}

\subsection{Results of Adaptive Sampling}
\label{sec:adaptiveres}
In order to test our QASS prototype, several functional toy theories for TBR were developed as alternatives to the expensive MC model. By far the most robust of these was the following sinusoidal theory with adjustable wavenumber parameter $n$:

@@

plotted in~\cref{fig:sintoy} for $n=1$ and two continuous parameters $C$.
ANNs trained on this model demonstrated similar performance to those on the expensive
MC model. QASS performance was verified by training a $\text{1h3f}(256)$ ANN on
the sinusoidal theory for varied quantities of initial, incremental, and MCMC
candidate samples. Although the scope of this project did not include thorough
searches of this hyperparameter domain, sufficient runs were made to identify
some likely trends.

An increase in initial samples with increment held constant had a strong impact
on final surrogate precision, an early confirmation of basic functionality. An
increase in MCMC candidate samples was seen to have a positive but very weak
effect on final surrogate precision, suggesting that the runtime of MCMC on each
iteration can be limited for increased efficiency. The most complex dynamics
arose with the adjustment of sample increment, shown in~\cref{fig:qassincr}. For
each tested initial sample quantity $N$, the optimal number of step samples was seen to be well-approximated by $\sqrt{N}$. The plotted error trends suggest that
incremental samples larger than this optimum give slower model improvement on both the training and evaluation sets, and a larger minimum error on the evaluation set. This performance distinction is predicted to be even more significant when trained on the expensive MC model, where the number of sample evaluations will serve as the primary bottleneck for computation time.
\begin{figure}[h!]
    \centering
    \vspace{-1ex}
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=1.1\linewidth]{fig6a_qassincrsamp.pdf}
    \end{subfigure}%
    \hfill%
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=1.1\linewidth]{fig6b_qassincrtime.pdf}
    \end{subfigure}
    \caption{QASS absolute training error over total sample quantity (left) and number of iterations (right). MAE represents surrogate error on the adaptively-sampled training/test set, and E\_MAE on the independent evaluation sets.}
    \label{fig:qassincr}
    \vspace{-1ex}
\end{figure}

@@

\begin{figure}[h]
\centering
\begin{minipage}{.48\textwidth}
  \centering
  \includegraphics[width=1.1\linewidth]{fig7_qasssampling.pdf}
  \captionof{figure}{Absolute training error for QASS, baseline scheme, and mixed scheme.}
  \label{fig:qasssampling}
\end{minipage}%
\hfill%
\begin{minipage}{.48\textwidth}
  \centering
  \includegraphics[width=1.1\linewidth]{fig8_qasssampling100k.pdf}
  \captionof{figure}{Absolute training error for QASS and baseline scheme, with 100k initial samples.}
  \label{fig:qasssampling100k}
\end{minipage}
\end{figure}
